### 2025-01-30 ESC195 Lecture Notes/main.tex
none

Evaluating the integral:
\begin{align*}
    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
    &= \lim_{b \to \infty} -e^{-b} + 1 \\
    &= 1
\end{align*}
\end{example}
\subsection{Convergence Test}
\paragraph{Convergence} An improper integral converges if the limit of the integral exists. 
\paragraph{Comparison Test} We can compare the integral to another integral to determine convergence or divergence.
\paragraph{Spliting the Integral} We can split the integral into two parts to determine convergence. We split it into two limits on the left and right, for example:
$$ \int_a^b f(x)\, dx = \lim_{t \to c^-} \int_a^t f(x) \, dx + \lim_{t \to c^+} \int_t^b f(x) \, dx $$
\section{Hyperbolic Trigonometric Functions}
\begin{definition}[Hyperbolic Sine]
    The hyperbolic sine function is defined as:
    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
\end{definition}
\begin{definition}[Hyperbolic Cosine]
    The hyperbolic cosine function is defined as:
    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
\end{definition}
\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
\begin{align}
    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
    \frac{d}{dx} \cosh(x) &= \sinh(x).
\end{align}
\paragraph{Identities} The hyperbolic functions satisfy the following identities:
\begin{align}
    \cosh^2(x) + \sinh^2(x) &= 1, \\
    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
    \sinh(2x) &= 2\sinh(x)\cosh(x).
\end{align}
\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$. Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g. $x = \cosh(t)$, $y = \sinh(t)$).
\paragraph{Area} The area of a sector of the hyperbola is given by:
\begin{equation}
    A = t/2,
\end{equation} 
where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
\begin{figure}[h!]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Draw the axes
            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};

@@

    \item \textbf{Limascons}: $r = a + b\sin \theta$ (heart with a hole)

@@

>>>>

Evaluating the integral:
\begin{align*}
    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
    &= \lim_{b \to \infty} -e^{-b} + 1 \\
    &= 1
\end{align*}
\end{example}
\subsection{Convergence Test}
\paragraph{Convergence} An improper integral converges if the limit of the integral exists. 
\paragraph{Comparison Test} We can compare the integral to another integral to determine convergence or divergence.
\paragraph{Spliting the Integral} We can split the integral into two parts to determine convergence. We split it into two limits on the left and right, for example:
$$ \int_a^b f(x)\, dx = \lim_{t \to c^-} \int_a^t f(x) \, dx + \lim_{t \to c^+} \int_t^b f(x) \, dx $$
\section{Hyperbolic Trigonometric Functions}
\begin{definition}[Hyperbolic Sine]
    The hyperbolic sine function is defined as:
    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
\end{definition}
\begin{definition}[Hyperbolic Cosine]
    The hyperbolic cosine function is defined as:
    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
\end{definition}
\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
\begin{align}
    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
    \frac{d}{dx} \cosh(x) &= \sinh(x).
\end{align}
\paragraph{Identities} The hyperbolic functions satisfy the following identities:
\begin{align}
    \cosh^2(x) - \sinh^2(x) &= 1, \\
    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
    \sinh(2x) &= 2\sinh(x)\cosh(x).
\end{align}
\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$. Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g. $x = \cosh(t)$, $y = \sinh(t)$).
\paragraph{Area} The area of a sector of the hyperbola is given by:
\begin{equation}
    A = t/2,
\end{equation} 
where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
\begin{figure}[h!]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Draw the axes
            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};

@@

    \item \textbf{Limasçons}: $r = a + b\sin \theta$ (heart with a hole)

@@

\paragraph{} Below are some examples of polar curves:
\begin{figure}[h!]
    \begin{minipage}{0.48\textwidth}
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, red] plot ({(1+cos(\x))*cos(\x)}, {(1+cos(\x))*sin(\x)});
            \node at (0,-3) {Cardioid: $r = 1 + \cos \theta$};
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, blue] plot ({cos(\x)*cos(\x)}, {cos(\x)*sin(\x)});
            \node at (0,-3) {Circle: $r = \cos \theta$};
    \end{minipage}

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, green] plot ({(1+0.5*sin(\x))*cos(\x)}, {(1+0.5*sin(\x))*sin(\x)});
            \node at (0,-3) {Limaçon: $r = 1 + 0.5\sin \theta$};
        \end{tikzpicture}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, orange] plot ({sqrt(abs(sin(2*\x)))*cos(\x)}, {sqrt(abs(sin(2*\x)))*sin(\x)});
            \node at (0,-3) {Lemniscate: $r^2 = \sin 2\theta$};
        \end{tikzpicture}
    \end{minipage}
\end{figure}

\begin{figure}[h!]
    \begin{minipage}{0.48\textwidth}
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, purple] plot ({cos(3*\x)*cos(\x)}, {cos(3*\x)*sin(\x)});
            \node at (0,-3) {Petal Curve: $r = \cos 3\theta$};
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, brown] plot ({sin(2*\x)*cos(\x)}, {sin(2*\x)*sin(\x)});
            \node at (0,-3) {Petal Curve: $r = \sin 2\theta$};
    \end{minipage}
\newpage
\paragraph{} \label{sec:polar} A detailed list of polar curves and their properties could be found in Appendix \ref{app:polarCurves}. 
